Methodology for Assessing the State of the Practice for Domain X

### 2021-10-20 StateOfPractice/Methodology/Methodology.tex
Revisions up to, but not including, Section 10 (Measure using measurement template)
Edits to measurement template section

\begin{enumerate}
\item Scope: Software is removed by narrowing what functionality is considered
  to be within the scope of the domain.
\item Usage: Software packages are eliminated if their installation procedure is
  not clear and easy to follow.
\item Age: The older software packages (age being measured by the last date when
  a change was made) are eliminated, except in the cases where an older software
  package appears to be highly recommended and currently in use.
\end{enumerate}

@@

@@

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=1.0\textwidth]{measurement_template}
    \caption{Top of Measurement Template}
    \label{measurement_template_image}
  \end{center}
\end{figure}

The Measurement Template is found in Appendix~\ref{SecGradingTemplate}.  This
template is used to track measurements and quality scores for all of the
software packages in the domain. For each software package, we fill out one
column of the template. This process can take between 1 to 4 hours for each
package.  Project developers can be contacted for help regarding installation,
if necessary, but a cap of about 2 hours should be imposed on the entire
installation process, to keep the overall measurement time feasible.

Most of the data to be collected should be straightforward from reviewing the
measurement template.  However, in some cases some extra guidance is necessary
to eliminate ambiguity, as follows:

\begin{enumerate}
\item Initial release date: Mark the release year if an exact date is not
  available.
\item Publications about the software: A list of publications can be found
  directly on the website of some software packages. For others use Google
  Scholar or a similar index.
\item Is there evidence that performance was considered?: Search the software
  artifacts for any mention of speed, storage, throughput, performance
  optimization, parallelism, multi-core processing, or similar
  considerations. The search function on GitHub can help.
\item Getting started tutorial: Sometimes this is found within another artifact,
  like the user manual.
\item Continuous integration: Search the software artifacts for any mention of
  continuous integration. The search function on GitHub can help.
\end{enumerate}

For completing the template, the following steps should be followed:

\begin{enumerate}
\item Gather the summary information into the top section of the document
  (Figure \ref{measurement_template_image})
\item Using the GitStats tool that is described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics -
  GitStats section found near the bottom of the document
\item Using the SCC tool that is also described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics - SCC
  section found near the bottom of the document
\item If the software package is found on git, gather the measurements for the
  Repo Metrics - GitHub section found near the bottom of the document
\item Review installation documentation and attempt to install the software
  package on a virtual machine
\item Gather the measurements for installability
\item Gather the measurements for correctness and verifiability
\item Gather the measurements for surface reliability
\item Gather the measurements for surface robustness
\item Gather the measurements for surface usability
\item Gather the measurements for maintainability
\item Gather the measurements for reusability
\item Gather the measurements for surface understandability
\item Gather the measurements for visibility and transparency
\item Assign a score out of ten for each quality. The score can be measured
  using the Measurement Template Impression Calculator, found in
  Appendix~\ref{SecImpressionCalculator}. For each quality measurement, the file
  indicates the appropriate score to assign the measurement based on possible
  measurement values.
\end{enumerate}

>>>>

\begin{enumerate}
\item Scope: Software is removed by narrowing what functionality is considered
  to be within the scope of the domain.
\item Usage: Software packages are eliminated if their installation procedure is
  not clear and easy to follow.
\item Age: The older software packages (age being measured by the last date when
  a change was made) are eliminated, except in the cases where an older software
  package appears to be highly recommended and currently in use.  (The Domain
  Expert can be consulted on this question, if necessary.)
\end{enumerate}

@@

\noindent The time frame of 18 months was selected as the separating point
between alive and dead projects because this is the usual timeframe for
operating system updates.

@@

The Measurement Template is found in Appendix~\ref{SecGradingTemplate}.  This
template is used to track measurements and quality scores for all of the
software packages in the domain. For each software package, we fill out one
column of the template. This process can take between 1 to 4 hours for each
package.  Project developers can be contacted for help regarding installation,
if necessary, but a cap of about 2 hours should be imposed on the entire
installation process, to keep the overall measurement time feasible.  To save
time, a
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}{blank
measurement template spreadsheet} has been prepared.  An excerpt of the
spreadsheet is shown in Figure~\ref{measurement_template_image}.

\begin{figure}[h!]
	\begin{center}
	  \includegraphics[width=1.0\textwidth]{measurement_template}
	  \caption{Excerpt of the Top of the Measurement Template}
	  \label{measurement_template_image}
	\end{center}
  \end{figure}

The full template consists of 108 questions categorized under 9 qualities.  The
questions were designed to be unambiguous, quantifiable and measurable with
limited time and domain knowledge. The measures are grouped under headings for
each quality, and one for summary information. The summary information (shown in
Figure~\ref{measurement_template_image}) is the first section of the template.
This section summarizes general information, such as the software name, number
of developers, etc.  We follow the definitions given by
\citet{GewaltigAndCannon2012} for the software categories.  Public means software
intended for public use.  Private means software aimed only at a specific
group, while the concept category is used for software written simply to demonstrate
algorithms or concepts. The three categories of development models are: open
source, where source code is freely available under an open source license;
free-ware, where a binary or executable is provided for free; and, commercial,
where the user must pay for the software product.

Following the summary section to measure 9 qualities: 
\begin{inparaenum}
	\item installability;
	\item correctness and verifiability;
	\item surface reliability;
	\item surface robustness;
	\item surface usability;
	\item maintainability;
	\item reusability;
	\item surface understandability; and,
	\item visibility/transparency. \end}inparaenum} Definitions of these
qualities are available in a working document on software quality.  Explain the
use of the word surface.

Most of the data to be collected should be straightforward from reviewing the
measurement template.  However, in a few cases extra guidance is necessary
to eliminate ambiguity, as follows:

\begin{enumerate}
\item Initial release date: Mark the release year if an exact date is not
  available.
\item Publications about the software: A list of publications can be found
  directly on the website of some software packages. For others use Google
  Scholar or a similar index.
\item Is there evidence that performance was considered?: Search the software
  artifacts for any mention of speed, storage, throughput, performance
  optimization, parallelism, multi-core processing, or similar
  considerations. The search function on GitHub can help.
\item Getting started tutorial: Sometimes this is found within another artifact,
  like the user manual.
\item Continuous integration: Search the software artifacts for any mention of
  continuous integration. The search function on GitHub can help.  In some
  cases, \texttt{yaml} files will provide a hint that continuous integration is
  employed.
\end{enumerate}

\noindent To fill-in the spreadsheet template, the following steps should be followed:

\begin{enumerate}
\item Gather the summary information into the top section of the document
  (Figure \ref{measurement_template_image}).
\item Using the GitStats tool that is described in
  Section~\ref{SecEmpiricalMeasures} gather the measurements for the Repo
  Metrics - GitStats section found near the bottom of the spreadsheet.
\item Using the SCC tool that is also described in
  Section~\ref{SecEmpiricalMeasures} gather the measurements for the Repo
  Metrics -- SCC section found near the bottom of the spreadsheet.
\item If the software package is found on git, gather the measurements for the
  Repo Metrics - the GitHub section found near the bottom of the spreadsheet.
\item Review installation documentation and attempt to install the software
  package on a virtual machine.
\item Gather the measurements for installability
\item Gather the measurements for correctness and verifiability
\item Gather the measurements for surface reliability
\item Gather the measurements for surface robustness
\item Gather the measurements for surface usability
\item Gather the measurements for maintainability
\item Gather the measurements for reusability
\item Gather the measurements for surface understandability
\item Gather the measurements for visibility and transparency
\item Assign a score out of ten for each quality. The score can be measured
  using the Measurement Template Impression Calculator, found in
  Appendix~\ref{SecImpressionCalculator}. For each quality measurement, the file
  indicates the appropriate score to assign the measurement based on possible
  measurement values.
\end{enumerate}

As in \citet{SmithEtAl2016}, Virtual machines (VMs) are used to provide an
optimal testing environments for each MGMP software product. VMs were used
because it is easier to start with a fresh environment without hav- ing to worry
about existing libraries and conflicts. Moreover, when the tests are complete
the VM can be deleted, without any impact on the host operating system. The most
significant advantage of using VMs is to level the playing field. Every software
install starts from a clean slate, which removes ``works-on-my-computer``
errors. In the grading data the details for each VM are noted, including
hypervisor and operating system version.
